---
title: 'Lecture 15: Multivariate Longitudinal Multilevel Models'
author: "Bayesian Psychometric Modeling"
output: html_document
---

```{r setup}
# Install/Load Packages ===============================================================================================

if (!require(ggplot2)) install.packages("ggplot2")
library(ggplot2)

if (!require(R2jags)) install.packages("R2jags")
library(R2jags)

if (!require(mcmcplots)) install.packages("mcmcplots")
library(mcmcplots)

```

Today's data and lecture inspiration come from Lesa Hoffman (http://www.lesahoffman.com) and her book Longitudinal Analysis: Modeling Within-Person Fluctuation and Change (http://www.pilesofvariance.com/). We will be examining data from Chapter 9: Time-Varying Predictors in Models of Change.

```{r dataimport}
Ch9Data = read.csv(file = "Chapter9.csv", header = TRUE)

# check out variable names
names(Ch9Data)

# check for missing data
apply(X = Ch9Data, MARGIN = 2, FUN = function(x) return(length(which(is.na(x)))))

```
From Lesa's Chapter 9 Handout:

>These simulated data are from Hoffman (2015) chapter 9, and include 200 girls measured approximately annually from ages 12–18 (time 0 = age 18) on their risky behavior (the outcome, a sum ranging from 10 to 50) and the extent to which their mothers monitored their activities (the time-varying predictor, a mean ranging from 1 to 5, centered at 3). A time-invariant predictor of the conservativeness of mothers’ attitudes about the smoking and drinking (a mean ranging from 1 to 5, centered at 4) was also collected at the age 12 occasion. Here are the individual growth trajectories for risky behavior and monitoring:

```{r spagplot}

plot1 = ggplot(data = Ch9Data, aes(x = age18+18, y = risky, group = PersonID))

## simple spaghetti plot
plot1 + geom_line() + stat_smooth(aes(group = 1)) + stat_summary(aes(group = 1),
    geom = "point", fun.y = mean, shape = 17, size = 3)

plot2 = ggplot(data = Ch9Data, aes(x = age18+18, y = mon3+3, group = PersonID))

## simple spaghetti plot
plot2 + geom_line() + stat_smooth(aes(group = 1)) + stat_summary(aes(group = 1),
    geom = "point", fun.y = mean, shape = 17, size = 3)

plot3 = ggplot(data = Ch9Data, aes(x = mon3+3, y = risky, group = PersonID))

## simple spaghetti plot
plot3 + geom_line() + stat_smooth(aes(group = 1)) + stat_summary(aes(group = 1),
    geom = "point", fun.y = mean, shape = 17, size = 3)
```

From p.1 of Lesa's handout (pulled 28 April, 2019)

> The best-fitting unconditional longitudinal models included fixed quadratic and random linear effects of age for risky behavior, but a random linear effect of
age for monitoring (although the fixed linear age slope was nonsignificant). In addition, mother’s attitudes significantly predicted the intercept and linear age slope for risky behavior, but did not significantly predict monitoring.

> Chapter 9 began with person-mean-centering and baseline-centering of monitoring of a time-varying predictor of risky behavior. Both were shown to be inadequate because they do not properly distinguish the intercept, linear age slope, and residual variance contained in the monitoring predictor, each of which could potentially relate to those of risky behavior. So the purpose of this example is to demonstrate alternative software methods of estimating models of multivariate change so that you can decide what approach (software and syntax combination) will be most optimal for your own data.

## Model 1: Undirected Multivariate Longitudinal Model

We will adapt Lesa's multilevel notation as we have a multivariate multilevel model. Most of the notation will remain the same, but the left-hand side of the Level 1 equation chagnes:

##### Level 1

$Monitor_{ti} = \beta_{0iM} +\beta_{1iM}\left(Age_{ti} -18\right) + e_{tiM}$

$Risky_{ti} = \beta_{0iR} +\beta_{1iR}\left(Age_{ti} -18\right) +\beta_{2iR}\left(Age_{ti} -18\right)^2 + e_{tiR}$

##### Level 2 (Monitor):

$\beta_{0iM} = \gamma_{00M} + U_{0iM}$

$\beta_{1iM} = \gamma_{10M} + U_{1iM}$ 

##### Level 2 (Risky):                                    

$\beta_{0iR} = \gamma_{00R} + \gamma_{01R}\left(Attitudes12_i -4 \right) + U_{0iR}$

$\beta_{1iR} = \gamma_{10R} + \gamma_{11R}\left(Attitudes12_i -4 \right) + U_{1iR}$

$\beta_{2iR} = \gamma_{20R}$

Additionally:

$$\left[ e_{tiR}, e_{tiM}\right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{R}\right),$$

where all unique elements of $\boldsymbol{R}$ are estimated (SAS' TYPE=UN structure), and:

$$\left[ U_{0iM}, U_{0iR}, U_{1iM}, U_{1iR} \right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{G}\right),$$
where all unique elements of $\boldsymbol{G}$ are estimated (SAS' TYPE=UN structure). 

We will need to create the two composite models for the dependent variables so as to code these into JAGS:

##### Composite Model for Monitor

$Monitor_{ti} = \left( \gamma_{00M} + U_{0iM}\right) + \left( \gamma_{10M} + U_{1iM}\right)\left(Age_{ti} -18\right) + e_{tiM}$

##### Composite Model for Risky
<!-- \left( \right) -->

$Risky_{ti} = \left( \gamma_{00R} + \gamma_{01R}\left(Attitudes12_i -4 \right) + U_{0iR} \right) + \left( \beta_{1iR} = \gamma_{10R} + \gamma_{11R}\left(Attitudes12_i -4 \right) + U_{1iR} \right)\left(Age_{ti} -18\right) + \left( \gamma_{20R}\right)\left(Age_{ti} -18\right)^2 + e_{tiR}$

Next, we need to create terms that JAGS will use to loop over during the data likelihood. Specifically, the number of level-2 observations (```N```), the number of level-1 observations for each level-2 observation (```PersonObs```), and the rows of the original data each level-1 observation within each level-2 observation occupies (```ObsRow```). We will also put both DVs into a single matrix (```X```).

```{r ModelSetup}

# need total number level-2 observations
N = length(table(Ch9Data$PersonID))

# need number of level-1 observations per level-2 observation
PersonObs = unlist(lapply(X = 1:N, FUN = function(x) return(length(which(Ch9Data$PersonID == x)))))

# format data for import into JAGS: need index for which rows of data belong to which time[person]
ObsRow = sapply(X = 1:N, FUN = function(x) return(which(Ch9Data$PersonID == x)), simplify = TRUE)

# create data matrix
X = cbind(Ch9Data$risky, Ch9Data$mon3)
```

Here, we will use very uninformative priors for all model parameters:

```{r Model1, cache=TRUE}
# create prior values for Wishart distributions for G and R
R0  = diag(2)
Rdf = 2

G0  = diag(4)
Gdf = 4

model01.function = function(){
  
  for (person in 1:N){
    for (time in 1:PersonObs[person]){
      
      # model for Risky
      meanVec[ObsRow[time, person],1] <- 
        (gamma00.risky + gamma01.risky*att4[ObsRow[time, person]] + U[person,1]) +                             # Risky Intercept
        (gamma10.risky + gamma11.risky*att4[ObsRow[time, person]] + U[person,3])*age18[ObsRow[time, person]] + # Risky Linear Age Slope
        (gamma20.risky)*age18[ObsRow[time, person]]^2                                                          # Risky Quadratic Age Slope
      
      # model for Monitor
      meanVec[ObsRow[time, person],2] <- 
        (gamma00.monitor + U[person, 2]) +                                                                    # Monitor Intercept
        (gamma10.monitor + U[person, 4])*age18[ObsRow[time, person]]                                          # Monitor Linear Age Slope
      
      X[ObsRow[time,person], 1:2] ~ dmnorm(meanVec[ObsRow[time,person], 1:2], R.inv[1:2, 1:2])
    }
    
  }
  # prior distribution for random effects
  for (person in 1:N){
    U[person, 1:4] ~ dmnorm(meanU[1:4], G.inv[1:4,1:4])
  }
  
  # hyper priors for U
  for (i in 1:4){
    meanU[i] <- 0 # zero mean
  }
  
  G.inv[1:4, 1:4] ~ dwish(G0[1:4, 1:4], Gdf)
  R.inv[1:2, 1:2] ~ dwish(R0[1:2, 1:2], Rdf)
  
  G[1:4, 1:4] <- inverse(G.inv[1:4, 1:4])
  R[1:2, 1:2] <- inverse(R.inv[1:2, 1:2])
  
  gamma00.risky   ~ dnorm(0, 0.0001)
  gamma01.risky   ~ dnorm(0, 0.0001)
  gamma10.risky   ~ dnorm(0, 0.0001)
  gamma11.risky   ~ dnorm(0, 0.0001)
  gamma20.risky   ~ dnorm(0, 0.0001)
  gamma00.monitor ~ dnorm(0, 0.0001)
  gamma10.monitor ~ dnorm(0, 0.0001) 
}

model01.data = list(
  N = N,
  X = X,
  PersonObs = PersonObs,
  ObsRow = ObsRow,
  R0 = R0,
  Rdf = Rdf,
  G0 = G0,
  Gdf = Gdf,
  att4 = Ch9Data$att4,
  age18 = Ch9Data$age18
)

model01.parameters = c("G", "R", "gamma00.risky", "gamma01.risky", "gamma10.risky", "gamma11.risky",
                       "gamma20.risky", "gamma00.monitor", "gamma10.monitor")

model01.seed = 27042019

model01.r2jags =  jags.parallel(
  data = model01.data,
  parameters.to.save = model01.parameters,
  model.file = model01.function,
  n.chains = 4,
  n.iter = 15000,
  n.thin = 1,
  n.burnin = 3000,
  jags.seed = model01.seed
)

model01.r2jags
traplot(model01.r2jags, parms = "R")

# R2jags is acting strange with n.burnin...and upon inspection, it is using burnin to adapt the algorithm.

model01a = window(x = as.mcmc(model01.r2jags), start = 2001, end = 12000)

model01a.results = cbind(round(summary(model01a)$statistics, 4), round(summary(model01a)$quantiles, 4), gelman.diag(model01a, multivariate = FALSE)[[1]][,1])
model01a.results
```


## Model 2: Undirected Multivariate Longitudinal Model

Next, we add Monitor to the prediction of Risky, using the entire value of Monitor in the prediction. Note, this corresponds to Mplus' ML estimator for this model. Again, will adapt Lesa's multilevel notation as we have a multivariate multilevel model. Most of the notation will remain the same, but the left-hand side of the Level 1 equation chagnes:

##### Level 1

$Monitor_{ti} = \beta_{0iM} +\beta_{1iM}\left(Age_{ti} -18\right) + e_{tiM}$

$Risky_{ti} = \beta_{0iR} +\beta_{1iR}\left(Age_{ti} -18\right) +\beta_{2iR}\left(Age_{ti} -18\right)^2 + \beta_{3iR}\left(Monitor_{ti}\right) e_{tiR}$

##### Level 2 (Monitor):

$\beta_{0iM} = \gamma_{00M} + U_{0iM}$

$\beta_{1iM} = \gamma_{10M} + U_{1iM}$ 

##### Level 2 (Risky):                                    

$\beta_{0iR} = \gamma_{00R} + \gamma_{01R}\left(Attitudes12_i -4 \right) + \gamma_{02R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{03M}\left(\gamma_{10M} + U_{1iM} \right) + U_{0iR}$

$\beta_{1iR} = \gamma_{10R} + \gamma_{11R}\left(Attitudes12_i -4 \right) + \gamma_{12R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{13M}\left(\gamma_{10M} + U_{1iM} \right)  + U_{1iR}$

$\beta_{2iR} = \gamma_{20R}$

$\beta_{3iR} = \gamma_{30R}$

Now, a few things are different:

$$\left[ e_{tiR}, e_{tiM}\right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{R}\right),$$

where all unique diagonal elements of $\boldsymbol{R}$ are estimated and the off-diagonal elements are set to zero (SAS' TYPE=VC structure), and:

$$\left[ U_{0iM}, U_{1iM}\right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{G}_M\right),$$

where all unique elements of $\boldsymbol{G}_M$ are estimated (SAS' TYPE=UN structure), and 

$$\left[ U_{0iR}, U_{1iR} \right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{G}_R\right),$$
where all unique elements of $\boldsymbol{G}_R$ are estimated (SAS' TYPE=UN structure). 

We will need to create the two composite models for the dependent variables so as to code these into JAGS:

##### Composite Model for Monitor

$Monitor_{ti} = \left( \gamma_{00M} + U_{0iM}\right) + \left( \gamma_{10M} + U_{1iM}\right)\left(Age_{ti} -18\right) + e_{tiM}$

##### Composite Model for Risky
<!-- \left( \right) -->

$Risky_{ti} = \left( \gamma_{00R} + \gamma_{01R}\left(Attitudes12_i -4 \right) + \gamma_{02R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{03M}\left(\gamma_{10M} + U_{1iM} \right) + U_{0iR} \right) + \left( \gamma_{10R} + \gamma_{11R}\left(Attitudes12_i -4 \right) + \gamma_{12R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{13M}\left(\gamma_{10M} + U_{1iM} \right)  + U_{1iR}\right)\left(Age_{ti} -18\right) + \left( \gamma_{20R}\right)\left(Age_{ti} -18\right)^2 + \left( \gamma_{30R} \right)Monitor_{ti} +e_{tiR}$

Here, we will break up the level-1 residuals to estimate with ```dnorm()```, which is equivalent to setting the covariance to zero.

```{r Model2, cache=TRUE}

# setting priors for variances -- keeping all terms independent as additional model parameters do the same job

# risky R matrix
sigma2.inv.risky.sse0 = .5*var(Ch9Data$risky)
sigma2.inv.risky.df0 = 1
    
sigma2.inv.risky.alpha0 = sigma2.inv.risky.df0/2
sigma2.inv.risky.beta0 = (sigma2.inv.risky.df0*sigma2.inv.risky.sse0)/2

# risky G matrix
G0risky  = c(.4*var(Ch9Data$risky), .1*var(Ch9Data$risky))*diag(2)
G0risky.df = 2

# monitor R matrix
sigma2.inv.monitor.sse0 = .5*var(Ch9Data$mon3)
sigma2.inv.monitor.df0 = 1
    
sigma2.inv.monitor.alpha0 = sigma2.inv.monitor.df0/2
sigma2.inv.monitor.beta0 = (sigma2.inv.monitor.df0*sigma2.inv.monitor.sse0)/2

# monitor G matrix
G0monitor = c(.4*var(Ch9Data$mon3), .1*var(Ch9Data$mon3))*diag(2)
G0monitor.df = 2

model02.function = function(){
  
  for (person in 1:N){
    for (time in 1:PersonObs[person]){
      
      # model for Risky
      meanVec[ObsRow[time, person],1] <- 
        (gamma00.risky + gamma01.risky*att4[ObsRow[time, person]] +                                    # Risky Intercept
           gamma02.risky*(gamma00.monitor + U.monitor[person,1]) + 
           gamma03.risky*(gamma10.monitor + U.monitor[person,2]) + U.risky[person,1]) +                             
        (gamma10.risky + gamma11.risky*att4[ObsRow[time, person]] +                                    # Risky Linear Age Slope
           gamma12.risky*(gamma00.monitor + U.monitor[person,1]) + 
           gamma13.risky*(gamma10.monitor + U.monitor[person,2]) +
           U.risky[person,2])*age18[ObsRow[time, person]] + 
        (gamma20.risky)*age18[ObsRow[time, person]]^2 +                                                # Risky Quadratic Age Slope
        gamma30.risky*Monitor[ObsRow[time,person]]                                                     # Risky Monitor Slope
      
      X[ObsRow[time,person], 1] ~ dnorm(meanVec[ObsRow[time,person], 1], sigma2.inv.risky)             # Univariate normal likelihood as R has zero covariance
      
      # model for Monitor
      meanVec[ObsRow[time, person],2] <-                                                               # Monitor Intercept
        (gamma00.monitor + U.monitor[person, 1]) +                                                                    
        (gamma10.monitor + U.monitor[person, 2])*age18[ObsRow[time, person]]                                   # Monitor Linear Age Slope
      
      X[ObsRow[time,person], 2] ~ dnorm(meanVec[ObsRow[time,person], 2], sigma2.inv.monitor)           # Univariate normal likelihood as R has zero covariance
    }
    
  }

  # prior distributions for R matrix variances
  sigma2.inv.risky       ~ dgamma(sigma2.inv.risky.alpha0,       sigma2.inv.risky.beta0)
  sigma2.inv.monitor     ~ dgamma(sigma2.inv.monitor.alpha0,     sigma2.inv.monitor.beta0)
  
  sigma2.risky      <- 1/sigma2.inv.risky
  sigma2.monitor    <- 1/sigma2.inv.monitor

    
  # prior distributions for random effects
  for (person in 1:N){
    U.risky[person, 1:2]   ~ dmnorm(U.risky.mean[1:2],   G.inv.risky[1:2,1:2])
    U.monitor[person, 1:2] ~ dmnorm(U.monitor.mean[1:2], G.inv.monitor[1:2,1:2])
  }
  
  # prior distributions for random effects G matrices
  G.inv.risky ~ dwish(G0risky, G0risky.df)
  G.risky <- inverse(G.inv.risky)
  
  G.inv.monitor ~ dwish(G0monitor, G0monitor.df)
  G.monitor <- inverse(G.inv.monitor)

  # prior distributions for fixed effects
  gamma00.risky   ~ dnorm(0, 0.0001)
  gamma01.risky   ~ dnorm(0, 0.0001)
  gamma10.risky   ~ dnorm(0, 0.0001)
  gamma11.risky   ~ dnorm(0, 0.0001)
  gamma20.risky   ~ dnorm(0, 0.0001)
  gamma02.risky   ~ dnorm(0, 0.0001)
  gamma03.risky   ~ dnorm(0, 0.0001)
  gamma12.risky   ~ dnorm(0, 0.0001)
  gamma13.risky   ~ dnorm(0, 0.0001)
  gamma30.risky   ~ dnorm(0, 0.0001)
  gamma00.monitor ~ dnorm(0, 0.0001)
  gamma10.monitor ~ dnorm(0, 0.0001)
  
}

model02.data = list(
  N = N,
  X = cbind(Ch9Data$risky, Ch9Data$mon3),
  PersonObs = PersonObs,
  ObsRow = ObsRow,
  att4 = Ch9Data$att4,
  age18 = Ch9Data$age18,
  Monitor = Ch9Data$mon3,
  sigma2.inv.risky.alpha0 = sigma2.inv.risky.alpha0,
  sigma2.inv.risky.beta0 = sigma2.inv.risky.beta0,
  sigma2.inv.monitor.alpha0 = sigma2.inv.monitor.alpha0,
  sigma2.inv.monitor.beta0 = sigma2.inv.monitor.beta0,
  G0monitor = G0monitor, 
  G0monitor.df = G0monitor.df,
  U.monitor.mean = rep(0,2),
  G0risky = G0risky,
  G0risky.df = G0risky.df,
  U.risky.mean = rep(0,2)
)

model02.parameters = c("gamma00.risky", "gamma01.risky", "gamma10.risky", "gamma11.risky",
                       "gamma20.risky", "gamma00.monitor", "gamma10.monitor", "gamma02.risky",
                       "gamma03.risky", "gamma12.risky", "gamma13.risky", "gamma30.risky",
                       "sigma2.risky", "sigma2.monitor", "G.risky", "G.monitor")

model02.seed = 27042019+1

model02.r2jags =  jags.parallel(
  data = model02.data,
  parameters.to.save = model02.parameters,
  model.file = model02.function,
  n.chains = 4,
  n.iter = 15000,
  n.thin = 1,
  n.burnin = 3000,
  jags.seed = model02.seed
)

model02.r2jags
traplot(model02.r2jags, parms = "gamma02.risky")

# R2jags is acting strange with n.burnin...and upon inspection, it is using burnin to adapt the algorithm.

model02a = window(x = as.mcmc(model02.r2jags), start = 2001, end = 12000)

model02a.results = cbind(round(summary(model02a)$statistics, 4), round(summary(model02a)$quantiles, 4), gelman.diag(model02a, multivariate = FALSE)[[1]][,1])
model02a.results

```



## Model 3: Undirected Multivariate Longitudinal Model with Residualized Monitor

Finally, we add just the residual version of Monitor to the prediction of Risky. Note, this corresponds to Mplus' Bayesian estimator for this model. Again, will adapt Lesa's multilevel notation as we have a multivariate multilevel model. 

##### Level 1

$Monitor_{ti} = \beta_{0iM} +\beta_{1iM}\left(Age_{ti} -18\right) + e_{tiM}$

$Risky_{ti} = \beta_{0iR} +\beta_{1iR}\left(Age_{ti} -18\right) +\beta_{2iR}\left(Age_{ti} -18\right)^2 + \beta_{3iR}\left(Monitor_{ti} -  \left(\left( \gamma_{00M} + U_{0iM}\right) + \left( \gamma_{10M} + U_{1iM}\right)\left(Age_{ti} -18\right) \right)\right) e_{tiR}$

##### Level 2 (Monitor):

$\beta_{0iM} = \gamma_{00M} + U_{0iM}$

$\beta_{1iM} = \gamma_{10M} + U_{1iM}$ 

##### Level 2 (Risky):                                    

$\beta_{0iR} = \gamma_{00R} + \gamma_{01R}\left(Attitudes12_i -4 \right) + \gamma_{02R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{03M}\left(\gamma_{10M} + U_{1iM} \right) + U_{0iR}$

$\beta_{1iR} = \gamma_{10R} + \gamma_{11R}\left(Attitudes12_i -4 \right) + \gamma_{12R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{13M}\left(\gamma_{10M} + U_{1iM} \right)  + U_{1iR}$

$\beta_{2iR} = \gamma_{20R}$

$\beta_{3iR} = \gamma_{30R}$

Now, a few things are different:

$$\left[ e_{tiR}, e_{tiM}\right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{R}\right),$$

where all unique diagonal elements of $\boldsymbol{R}$ are estimated and the off-diagonal elements are set to zero (SAS' TYPE=VC structure), and:

$$\left[ U_{0iM}, U_{1iM}\right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{G}_M\right),$$

where all unique elements of $\boldsymbol{G}_M$ are estimated (SAS' TYPE=UN structure), and 

$$\left[ U_{0iR}, U_{1iR} \right]^T \sim MVN\left( \boldsymbol{0}, \boldsymbol{G}_R\right),$$
where all unique elements of $\boldsymbol{G}_R$ are estimated (SAS' TYPE=UN structure). 

We will need to create the two composite models for the dependent variables so as to code these into JAGS:

##### Composite Model for Monitor

$Monitor_{ti} = \left( \gamma_{00M} + U_{0iM}\right) + \left( \gamma_{10M} + U_{1iM}\right)\left(Age_{ti} -18\right) + e_{tiM}$

##### Composite Model for Risky
<!-- \left( \right) -->

$Risky_{ti} = \left( \gamma_{00R} + \gamma_{01R}\left(Attitudes12_i -4 \right) + \gamma_{02R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{03M}\left(\gamma_{10M} + U_{1iM} \right) + U_{0iR} \right) + \left( \gamma_{10R} + \gamma_{11R}\left(Attitudes12_i -4 \right) + \gamma_{12R}\left(\gamma_{00M} + U_{0iM}\right) + \gamma_{13M}\left(\gamma_{10M} + U_{1iM} \right)  + U_{1iR}\right)\left(Age_{ti} -18\right) + \left( \gamma_{20R}\right)\left(Age_{ti} -18\right)^2 + \left( \gamma_{30R} \right) \left(Monitor_{ti} -  \left(\left( \gamma_{00M} + U_{0iM}\right) + \left( \gamma_{10M} + U_{1iM}\right)\left(Age_{ti} -18\right) \right)\right) +e_{tiR}$


```{r Model3, cache=TRUE}

# setting priors for variances -- keeping all terms independent as additional model parameters do the same job

# risky R matrix
sigma2.inv.risky.sse0 = .5*var(Ch9Data$risky)
sigma2.inv.risky.df0 = 1
    
sigma2.inv.risky.alpha0 = sigma2.inv.risky.df0/2
sigma2.inv.risky.beta0 = (sigma2.inv.risky.df0*sigma2.inv.risky.sse0)/2

# risky G matrix
G0risky  = c(.4*var(Ch9Data$risky), .1*var(Ch9Data$risky))*diag(2)
G0risky.df = 2

# monitor R matrix
sigma2.inv.monitor.sse0 = .5*var(Ch9Data$mon3)
sigma2.inv.monitor.df0 = 1
    
sigma2.inv.monitor.alpha0 = sigma2.inv.monitor.df0/2
sigma2.inv.monitor.beta0 = (sigma2.inv.monitor.df0*sigma2.inv.monitor.sse0)/2

# monitor G matrix
G0monitor = c(.4*var(Ch9Data$mon3), .1*var(Ch9Data$mon3))*diag(2)
G0monitor.df = 2

model03.function = function(){
  
  for (person in 1:N){
    for (time in 1:PersonObs[person]){
      
      # residualize monitor variable for prediction of risky
      residualMonitor[ObsRow[time, person]] <- Monitor[ObsRow[time,person]] - 
        ((gamma00.monitor + U.monitor[person, 1]) + (gamma10.monitor + U.monitor[person, 2])*age18[ObsRow[time, person]])
      
      # model for Risky
      meanVec[ObsRow[time, person],1] <- 
        (gamma00.risky + gamma01.risky*att4[ObsRow[time, person]] +                                    # Risky Intercept
           gamma02.risky*(gamma00.monitor + U.monitor[person,1]) + 
           gamma03.risky*(gamma10.monitor + U.monitor[person,2]) + U.risky[person,1]) +                             
        (gamma10.risky + gamma11.risky*att4[ObsRow[time, person]] +                                    # Risky Linear Age Slope
           gamma12.risky*(gamma00.monitor + U.monitor[person,1]) + 
           gamma13.risky*(gamma10.monitor + U.monitor[person,2]) +
           U.risky[person,2])*age18[ObsRow[time, person]] + 
        (gamma20.risky)*age18[ObsRow[time, person]]^2 +                                                # Risky Quadratic Age Slope
        gamma30.risky*residualMonitor[ObsRow[time, person]]                                            # Risky Monitor Slope
      
      X[ObsRow[time,person], 1] ~ dnorm(meanVec[ObsRow[time,person], 1], sigma2.inv.risky)             # Univariate normal likelihood as R has zero covariance
      
      # model for Monitor
      meanVec[ObsRow[time, person],2] <-                                                               # Monitor Intercept
        (gamma00.monitor + U.monitor[person, 1]) +                                                                    
        (gamma10.monitor + U.monitor[person, 2])*age18[ObsRow[time, person]]                                   # Monitor Linear Age Slope
      
      X[ObsRow[time,person], 2] ~ dnorm(meanVec[ObsRow[time,person], 2], sigma2.inv.monitor)           # Univariate normal likelihood as R has zero covariance
    }
    
  }

  # prior distributions for R matrix variances
  sigma2.inv.risky       ~ dgamma(sigma2.inv.risky.alpha0,       sigma2.inv.risky.beta0)
  sigma2.inv.monitor     ~ dgamma(sigma2.inv.monitor.alpha0,     sigma2.inv.monitor.beta0)
  
  sigma2.risky      <- 1/sigma2.inv.risky
  sigma2.monitor    <- 1/sigma2.inv.monitor

    
  # prior distributions for random effects
  for (person in 1:N){
    U.risky[person, 1:2]   ~ dmnorm(U.risky.mean[1:2],   G.inv.risky[1:2,1:2])
    U.monitor[person, 1:2] ~ dmnorm(U.monitor.mean[1:2], G.inv.monitor[1:2,1:2])
  }
  
  # prior distributions for random effects G matrices
  G.inv.risky ~ dwish(G0risky, G0risky.df)
  G.risky <- inverse(G.inv.risky)
  
  G.inv.monitor ~ dwish(G0monitor, G0monitor.df)
  G.monitor <- inverse(G.inv.monitor)

  # prior distributions for fixed effects
  gamma00.risky   ~ dnorm(0, 0.0001)
  gamma01.risky   ~ dnorm(0, 0.0001)
  gamma10.risky   ~ dnorm(0, 0.0001)
  gamma11.risky   ~ dnorm(0, 0.0001)
  gamma20.risky   ~ dnorm(0, 0.0001)
  gamma02.risky   ~ dnorm(0, 0.01)
  gamma03.risky   ~ dnorm(0, 0.0001)
  gamma12.risky   ~ dnorm(0, 0.0001)
  gamma13.risky   ~ dnorm(0, 0.0001)
  gamma30.risky   ~ dnorm(0, 0.0001)
  gamma00.monitor ~ dnorm(0, 0.0001)
  gamma10.monitor ~ dnorm(0, 0.0001)
  
}

model03.data = list(
  N = N,
  X = cbind(Ch9Data$risky, Ch9Data$mon3),
  PersonObs = PersonObs,
  ObsRow = ObsRow,
  att4 = Ch9Data$att4,
  age18 = Ch9Data$age18,
  Monitor = Ch9Data$mon3,
  sigma2.inv.risky.alpha0 = sigma2.inv.risky.alpha0,
  sigma2.inv.risky.beta0 = sigma2.inv.risky.beta0,
  sigma2.inv.monitor.alpha0 = sigma2.inv.monitor.alpha0,
  sigma2.inv.monitor.beta0 = sigma2.inv.monitor.beta0,
  G0monitor = G0monitor, 
  G0monitor.df = G0monitor.df,
  U.monitor.mean = rep(0,2),
  G0risky = G0risky,
  G0risky.df = G0risky.df,
  U.risky.mean = rep(0,2)
)

model03.parameters = c("gamma00.risky", "gamma01.risky", "gamma10.risky", "gamma11.risky",
                       "gamma20.risky", "gamma00.monitor", "gamma10.monitor", "gamma02.risky",
                       "gamma03.risky", "gamma12.risky", "gamma13.risky", "gamma30.risky",
                       "sigma2.risky", "sigma2.monitor", "G.risky", "G.monitor")

model03.seed = 27042019+2

model03.r2jags =  jags.parallel(
  data = model03.data,
  parameters.to.save = model03.parameters,
  model.file = model03.function,
  n.chains = 4,
  n.iter = 15000,
  n.thin = 1,
  n.burnin = 3001,
  jags.seed = model03.seed
)

model03.r2jags
traplot(model03.r2jags, parms = "gamma02.risky")


model03a = window(x = as.mcmc(model03.r2jags), start = 2001, end = 12000)
model03a.results = cbind(round(summary(model03a)$statistics, 4), round(summary(model03a)$quantiles, 4), gelman.diag(model03a, multivariate = FALSE)[[1]][,1])

model23comp = cbind(model02a.results[,1:2], model03a.results[,1:2])
model23comp = cbind(model23comp, model23comp[,1] - model23comp[,3])
colnames(model23comp) = c("M2.PosteriorMean","M2.PosteriorSD", "M3.PosteriorMean","M3.PosteriorSD","M2-M3 Mean Dif")
model23comp
```

Notice the difference in two parameter values...ask Lesa Hoffman why that happens to be.
